---
title: "Manhattan Residential Condominiums"
author: "Niket"
date: "July 26, 2018"
output: rmarkdown::github_document
---

```{r}
manhtn18 <- read.csv(file.choose(), header = TRUE, skip = 1, na.strings = c("NA", "", " ", "."))
```

```{r}
# remove first empty row
manhtn18 <- manhtn18[-1,]
```


```{r}
str(manhtn18)
```

We should only include columns with Brooklyn listings without comparable rentals, which is 1 to 15 columns
```{r}
manhtn <- manhtn18[,1:15]
```

***
***

```{r}
summary(manhtn$Building.Class)
```
R2	CONDO; RESIDENTIAL UNIT IN WALK-UP BLDG.
R4	CONDO; RESIDENTIAL UNIT IN ELEVATOR BLDG.
R9	CO-OP WITHIN A CONDOMINIUM
RR	CONDO RENTALS


Elevator or walk-up (but walk-up must be lower than 3rd floor) and
Must be condominium (not co-op or condop) so we should drop R9 and Missing values(NA's)
```{r}
manhtn <- manhtn[!manhtn$Building.Class == "R9  -CONDOPS",]
manhtn <- manhtn[!is.na(manhtn$Building.Class),]
# manhtn_1 <- manhtn_1[!manhtn$Building.Class == "RR  -CONRENT",]
```

```{r}
summary(manhtn$Building.Class)
```

***
***

Market Value: less than $3,500,000

```{r}
manhtn$Full.Market.Value <- as.numeric(gsub(",", "", manhtn$Full.Market.Value)) #As the numerical values in Price column are comma seperated, remove comma

summary(manhtn$Full.Market.Value)
```

```{r}
manhtn <- manhtn[manhtn$Full.Market.Value < 3500000,]

summary(manhtn$Full.Market.Value)
```

***
***

Price per square foot: less than $2,000

```{r}
# manhtn$Market.Value.per.SqFt <- as.numeric(gsub(",","", manhtn$Market.Value.per.SqFt)) #As the numerical values in Price.per.sq.foot column are comma seperated, remove comma

summary(manhtn$Market.Value.per.SqFt)
```
All the data is under $2,000, but we can still clean for future purposes

```{r}
manhtn <- manhtn[manhtn$Market.Value.per.SqFt < 2000,]

summary(manhtn$Market.Value.per.SqFt)
```

***
***

Close by Location: Manhattan south of 96th Street

```{r}
manhtn$Address <- as.character(manhtn$Address)
```

Add NY to all the address
```{r}
manhtn$Address <- paste0(manhtn$Address, " NY")
```


```{r}
# install.packages("ggmap") # an R interface to the Google Maps API
```

```{r}
sum(is.na(manhtn)) #check for missing values
```


```{r}
# calculate distance between two place names
# we added NY to the address so that it does not pick location from another state
library(ggmap)
dist_96 <- mapdist(from = manhtn$Address, to = "Manhattan south of 96th Street")
```

```{r}
# fill distance from 96th street in our data
manhtn$Distance.to.96.Street <- dist_96$miles[match(manhtn$Address, dist_96$from)]
```

```{r}
# Looking at the distance
summary(manhtn$Distance.to.96.Street)
```


***
***

Rental yield of 2% or more = Net Income divided by Total Price

```{r}
# Rename BROOKLYN.Ã.Â.Â..CONDOMINIUMS.COMPARABLE.PROPERTIES.Ã.Â.Â..Net.Operating.Income to Net.Operating.Income
# colnames(manhtn_1)[colnames(manhtn_1)=="BROOKLYN.Ã.Â.Â..CONDOMINIUMS.COMPARABLE.PROPERTIES.Ã.Â.Â..Net.Operating.Income"] <- "Net.Operating.Income"
```

```{r}
manhtn$Net.Operating.Income <- as.numeric(gsub(",", "", manhtn$Net.Operating.Income)) #As the numerical values in Price.per.sq.foot column are comma seperated, remove comma

summary(manhtn$Net.Operating.Income)
```

```{r}
manhtn$Rental.Yield <- ((manhtn$Net.Operating.Income / manhtn$Full.Market.Value) * 100) 
```


```{r}
manhtn <- manhtn[manhtn$Rental.Yield >= 2,] # Rental yield of 2% or more

summary(manhtn$Rental.Yield)
```

***
***


Look for the closest locations(under 5 miles) order by Market Value
```{r}
distance_order <- manhtn[with(manhtn,order(Distance.to.96.Street)),]
distance_order <- distance_order[distance_order$Distance.to.96.Street < 5,]
```

```{r}
price_order <- distance_order[with(distance_order, order(Full.Market.Value)),]
price_order[1:30,]
```

***
***
